Accelerated convergence of convective simulations using boundary value problems

### 2018-02-13 tex/bvp_paper.tex
rough restructure of the intro; adds 7 sources and some discussion of various methods for converging solutions fast used historically; will come back to it tomorrow.

Prior studies of stratified convection in which a convective layer lies between stable layers
have used the knowledge of Mixing Length Theory (MLT) to adjust the initial thermodynamic structure of
the atmosphere to a state which is closer to the adiabat chosen by convection \cite{brandenburg&all2005}.
However, many modern studies of convection do not contain stable layers above and below the convection
zone, and the presence of hard boundaries and the boundary layers that they form make it difficult to
know the correct evolved adiabat \emph{a priori}.

Here we present a method for using simple boundary value problems (BVPs), 
along with information about the evolved flow fields,
to fast-forward the slow thermal evolution of convecting simulations.  
We run two sets of experiments: one in which we allow convective simulations to evolve for a
full thermal timescale before taking measurements, and another in which we employ a fast,
BVP technique which occurs on dynamical timescales. We compare these two sets of simulations to
show the validity of the BVP technique.  We use the BVP technique to run simulations
at high Ra, in the regime where running for thermal timescales becomes computationally intractable.

>>>>

The vast difference between convective and diffusive timescales has long plagued
numericists studying convection, and a vast number of approaches has been used to
study thermally converged solutions. ``Bootstrapping'' is the process of using
a converged solution at lower Ra as initial conditions for a simulation at a higher
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  This method is likely to run into hystersis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
For moderate-Ra solutions, a commonly-used tactic is to use initial conditions from
a simpler model of the full convective state.  For example, using the results of
a linear solve \cite{hurlburt&all1984} or using an axisymmetric solution for a full
3D cylinder of convection \cite{verzicco&camussi1997}. In certain systems, 
the physics of the evolved solution can be approximately known based on the initial
conditions, and there a better set of initial conditions can be solved for analytically
\cite{couston&all2017} or by using knowledge of Mixing Length Theory or other convective
theories to adjust the initial profile towards the proper adiabat \cite{brandenburg&all2005}.

Regardless, the most straightforward way of achieving a thermally converged solution
is by simply allowing a simulation to run for a thermal timescale, which some
studies do \cite{featherstone&hindman2016}.  Unfortunately, such a procedure is
\emph{expensive}, and state-of-the-art simulations can often only be reasonably run
for tens to hundreds of buoyancy times \cite{stevens&all2011}, much less the
thousands-to-millions of buoyancy timescales required to reach a thermal time
at high Ra.

One further method of adjusting the atmospheric profile is by using information about
the convective dynamics, as well as knowledge about energy balances and fluxes in
the evolved solution, to carefully adjust the mean thermodynamic profile towards
its proper solution.  Such methods have some history, \cite{hurlburt&all1986}, but
a clear presentation of such a method and a comparison of such a method to solutions
which were achieved through a long thermal rundown are lacking.
Here we present a simple 1D model that uses information about evolving flow fields
to accelerate the evolution of the thermal profile in convective simulations.
We run two sets of experiments: one in which we allow convective simulations to evolve for a
full thermal timescale before taking measurements, and another in which we employ a fast,
BVP technique which occurs on dynamical timescales. We compare these two sets of simulations to
show the validity of the BVP technique.  We use the BVP technique to run simulations
at high Ra, in the regime where running for thermal timescales becomes computationally intractable.
